AttCRISPR : a spacetime interpretable model for prediction of sgRNA on-target activity

### 2020-11-12 Paper/main.tex
11-12 paper

\abstract{\textbf{Motivation:} More and more Cas9 variants with higher specificities are developed to avoid the off-target effect, which bring a significant volume of experimental data. 
Conventional machine learning performs poorly on these datasets, while the models based on deep learning are often lack of interpretability, which makes it difficult for researchers to understand its decisions, 
not to mention to design high-activity sgRNA with the insights brought by the model. 
Moreover, neither of the deep learning based model with existing structure can not satisfy enough precision at such huge datasets.\\
\textbf{Results:} To overcome these problems, we design and implement AttCRISPR, an intrinsic interpretable method based on deep learning to predict the on-target activity. 
In this paper, we find that based on the preprocessing of the input base sequence, the deep learning models can be divided into two categories, 
methods in spatial domain with the input of encoded base sequence and in temporal domain with the input of embedded base sequence. 
Based on the categorization above, we assume that spatial and temporal method are heterogeneous, and use a full connection layer to stack them, which achieve a better performance. 
Our model was trained and tested on the biggest dataset, the same dataset as DeepHF used, as far as we know for performance evaluation. 
In addition, benefits from two attention modules, spatial and temporal attention modules, AttCRISPR has strong interpretability. 
Through them, we can understand the decisions made by AttCRISPR at both global and local levels without other post hoc explanations techniques. 
Further, interpretability AttCRISPR can be used to guide researcher to design sgRNA with higher activity, in order to reduce the experimental cost. 
AttCRISPR can compete with state-of-the-art method in these datasets, and achieves an average spearman value of 0.872, 0.867, 0.867 (corresponding to WT-SpCas9, eSpCas9(1.1), SpCas9-HF1) under tenfold shuffled validation. \\
\textbf{Availability:} The example code are available at \href{https://github.com/South-Walker/AttCRISPR}{https://github.com/South-Walker/AttCRISPR}\\
\textbf{Contact:} \href{zrjiang@cs.ecnu.edu.cn}{zrjiang@cs.ecnu.edu.cn}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
online.}

@@

Clustered regularly interspaced short palindromic repeats (CRISPR) / CRIPSR associated protein 9 (Cas9) systems are preferred over other biological research and human medicine technologies now, 
beacuse of it's efficiency, robustness and programmability. 
Cas9 nucleases can be directed by short guide RNAs (sgRNAs) to introduce site-specific DNA double-stranded breaks in target, 
so to enable editing site-specific regions within genome \citep{jinek2012a,cong2013multiplex,mali2013rnaguided}. 
CRISPR/Cas9, to a large extent, has developed genetic therapies at the cellular level, 
while there are still severe medical disadvantages even now which has greatly hindered the further clinical application of the CRISPR/Cas9 systems. 
One of these disadvantages is due to unexpected insertion and deletion caused by off-target effect \citep{rubeis2018risks,kang2016introducing,ishii2017reproductive,liang2015crispr/cas9-mediated}. 
To overcome this disadvantage, a solution is to engineer CRISPR / Cas9 with higher specificities. 
That's why more and more higher specificities Cas9 variants, such as enhanced SpCas9 (eSpCas9(1.1)), 
Cas9-High Fidelity (SpCas9-HF1) \citep{ishii2017reproductive,slaymaker2016rationally}, hyper-accurate Cas9 (HypaCas9) \citep{kleinstiver2016high-fidelity}, 
been developed and bring a significant volume of experimental data, which means that researchers have to face the difficulty of analyzing such huge and heterogeneous data.

@@

 In addition, existing analysis of the models interpretability are all at global level, and result is a general pattern in the datasets. 
For example, which position has a great impact on the final decision of the model and which position-dependent nucleotide has a positive impact on the activity. 
Lack of analysis at local level. 
For example, which structure causes the low activity of a certain nucleotide sequence and how to improve its activity with a few modification. 
In light of the above, we believe it is essential to develope a model which can not only match deep learning based model in performance, 
but also with strong interpretability that can be comparable to conventional machine learning method.

@@

Attention mechanism has demonstrated its power in NLP, Statistical Learning, Speech and Computer Vision. 
 It makes model tends to focus selectively on parts of the input, which is help in performing the task effectively. 
 Previous observation have shown that Cas9 preferentially binds sgRNAs containing purines but not pyrimidines \citep{wang2014genetic} and multiple thymine in the spacer impairing sgRNA activity \citep{wu2014genome-wide}, 
 that is to say some specific nucleotides and base position need more attention compared to others. 
 The above is the premise of introducing attention mechanism. Strictly speaking, we are not the first to bring attention mechanisms into this field. 
 The most similar approach to ours is the work based on transformer by \citeauthor{Liu2019}. 
 They use transformer, a components based on attention mechanism, instead of RNN to improve the ability of temporal feature extraction, 
 hence, enhance the performance of their model \citep{Liu2019,vaswani2017attention}. 
 In our work,the interpretability benefit from attention mechanism is more focused. 
 Our main contributions are as follows:\vspace*{1pt}
\begin{itemize}
    \item Present a novel deep-learning model, which can extract potential feature representation of sgRNA sequence in both spatial and temporal domain parallelly. 
    Finally, the ensemble learning method is used to combine the two to achieve better performance than current state-of-the-art models. 
    \item Introduce attention mechanism into our model. 
    As a result, it does not need post hoc explanations techniques based on input perturbation to explain itself. 
    It is intrinsic interpretable in both temporal and spatial domains.
    In the spatial domain it's at global level, while at local level in the temporal domain. 
    Thus, it is transformed from a black box to an intrinsically interpretable model with the performance of deep learning based model. 
    \item Throught ablation analysis and testing a series of possible network structures, 
    we find there are multiple components and strategy can improve the performance of our model and constructed AttCRISPR, 
    which could outperform current state-of-the-art tools on DeepHF dataset.\vspace*{1pt}
\end{itemize}

\section{Materials and methods}

@@

For encoding process, we use the complementary base to represent the original base in sgRNA, as others do. 
Further, we use one-hot encode strategy, which's is to say, we encoding each base in sgRNA into a four-dimensional vector 
(encode A,T,G,C into [1,0,0,0], [0,1,0,0], [0,0,1,0], [0,0,0,1], respectively), called one-hot vector. 
Then a sgRNA can be considered as a matrix $X_{oh}\in\mathbb{R}^{l\times4} $, named one-hot matrix (a little sparse, since an one-hot vector is zero in all but one dimension). 
We believe it is meaningful to regard $X_{oh}$ as a binary image, 
therefore, it is used as an input of convolutional neural network, which performs well in the image field.
Meanwhile, as mentioned above, one-hot matrix is a little sparse.

To facilitate the training process, we can mapping each one-hot vector into a dense real-valued high-dimensional space, which is called embedding. 
Concretely speaking, at the matrix level, the formula is as follows
\begin{equation}
X_e=X_{oh}E_m\label{eq:08}
\end{equation}
where $X_e$ named embedding matrix, $E_m\in\mathbb{R}^{4\times m}$ is a trainable transformational matrix, $m$ refers to the dimension of embedding space. 
We believe it is also meaningful to regard nucleotides in the sgRNA sequence as word, and the sgRNA sequence itself as a sentence, 
Guided by this belief $E_m$ is the word embedding matrix and $X_e$ is the sentence embedding in NLP.
Therefore, $X_e$ is used as an input of RNN (or its variant), which performs well in the NLP field.

Due to each element of $X_{oh}$ is interpretable (representing whether there is a corresponding nucleotide type at the corresponding location), 
we call $X_{oh}$ the spatial input, and the convolutional neural network work on $X_{oh}$ is method in spatial domain. 
In other hand, different from $X_{oh}$, $X_e$ can only be explained in the first dimension (representing the embedding vector of corresponding nucleotide type), 
and embedding vector is difficult for humans to understand. 
That's why we call $X_e$ the temporal input, and the RNN (or its variant) work on $X_e$ is method in temporal domain.

@@

Before introducing the neural network architecture of AttCRISPR, Let's define first-order preference and second-order preference for convenience. 
Taking a simple linear regression model as an example, for input $X$ in Equation~(\ref{eq:01}), $y$ in Equation~(\ref{eq:01}) is as following
 \begin{equation}
y = AX\label{eq:linear}
\end{equation}
where $A \in\mathbb{R}^d$, the total differential of $y$ in Equation~(\ref{eq:linear}) is as following
 \begin{equation}
dy = \sum^{d}_iA_idX_i\label{eq:02}
\end{equation}
where $A_i$ and $X_i$ denotes the $i$-th dimension of the vector $X$ and $A$, 
$A_i$ indicates how dramatically the function changes as $X_i$ changes in a neighborhood of $X$, 
in other word, the importance of $X_i$. 
That's why we'll call $A$ first-order preference in our paper. 
Specifically, we use a vector $A_i$ to build the first-order original preference at position $i$ within sgRNA sequence, 
and $X_i$ is an embeddedness of the $i$-th feature, then $A$ and $X$ are two matrix. 
Further, the final result can be weighted by a trainable non-negative weight vector $W\in\mathbb{R}^l$, as follow
\begin{equation}
y=W \cdot AX^{T}\label{eq:superlinear}
\end{equation}
then we define $\tilde{A}$ as the first-order combine preference matrix (or just first-order preference), which means $\tilde{A}$ can be expressed linearly by $A$ as follow
\begin{equation}
\tilde{A}=BA\label{eq:04}
\end{equation}
where the weight matrix $B\in\mathbb{R}^{l\times l}$ is learned through attention mechanism, 
which we'll call the second-order preference matrix in our paper due to its calculation is based on first-order preference, 
it can explain how a particular pattern containing two nucleotides affects the base sequence. 
Then the predicted value can be expressed as
\begin{equation}
y=W \cdot \tilde{A} {X_e}^{T}\label{eq:05}
\end{equation}

@@

As shown in Figure~\ref{fig:spatialmodule}, given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{array}{l}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{array}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other words, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, we train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.

@@

    Although it's not shown above, it is also worth noting that, a full connection layer with L2 constraints is used to generates a transformation matrix in order to ensure $\tilde{A}_i$ and ${X_e} _i$ are in the same vector space. 

@@

Then we use a simple full connection network to extract the indirect features, and call the output of fully connection network $y_{bio}$. 
As mentioned above, we assume that the method in spatial domain and in temporal domain are heterogeneous, which can satisfy the diversity premise of ensemble learning.
That's why we follow the stacking strategy, to integrate the methods in time domain and space domain. 
Specifically, the $y_{bio}$, the spatial output $y_{s}$ and the temporal output $y_{t}$ we got earlier are concatenated and then weighted averaging is performed through a full connection layer as follow
\begin{equation}
y=W[y_{bio};y_s;y_t]\label{eq:20}
\end{equation}
where, $y$ is the final prediction value of AttCRISPR, $W$ is the weight learned by the full connection network. 
In the actual implementation, we freeze the network in the spatial domain and temporal domain firstly, in order to make our network focused on learning the weight $W$.
Then the parameters of the entire network are adjusted in the fine tuning of AttCRISPR.

\subsection{Datasets and current prediction method}

The dataset we used for training, validation and testing is built by \citeauthor{wang2019optimized} \citep{wang2019optimized}. 
We extracted 55604, 58617, 56888 sgRNAs with activity (represented by insertion/deletion (indel)) for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively, from its source data. 
As far as we know, the best performance on this dataset is the DeepHF of \citeauthor{wang2019optimized} currently, which is based on deep learning and bidirectional LSTM architecture. 
DeepHF achieved Spearman correlation coefficients of 0.867, 0.862 and 0.860 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively.

Another tool worth noting is CRISPRpred(SEQ), which is based on conventional machine learning and need complicated human feature engineering exercise \citep{MuhammadRafid2020}. 
They also test their tool with this dataset and achieved Spearman correlation coefficients of 0.838, 0.830 and 0.821 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively (without any hyperparameter tuning).

\citeauthor{wang2019optimized} implemente a tenfold shuffled validation to evaluate the stability of the performance of DeepHF.
To be more specific, each set is shuffled and divided into three parts, 76.5\%, 15\% and 8.5\% of the relevant data was used as the training, test and validation set respectively in a single experiment. 
The experiment is repeated ten times with the results recorded and averaged finally.
In order to make the comparison apples to apples, we will follow this strategy to design experiments.

@@

\section{Results}
\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{baselinewithoutbiofeat.eps}}
    \caption{In the absence of hand-crafted biological features, performance of different algorithms for sgRNA activity prediction. 
    (a)-(c) The performance of Temporal AttCRISPR, Spatial AttCRISPR and Ensemble AttCRISPR. 
    The half-violin plots show the mean and distribution of the Spearman correlation coefficient between predicted and measured sgRNA activity scores over all tests. 
    (d)-(f) In the absence of hand-crafted biological features, the performance of all prediction methods in these datasets as far as we know. 
    The $mean \pm s.d.$ of the Spearman correlation coefficient between predicted and measured sgRNA activity scores are shown in the bar plots.}\label{fig:06}
\end{figure}
We designed experiments to address the following questions:
\begin{itemize}
    \item 
    In the absence of hand-crafted biological features, whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions made by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

@@

In the following sections, we will analyze the insight into activity of sgRNA bring through the attention mechanism at both global and local levels 
to validate the attention module in the AttCRISPR can help us to understand the decisions. 
\begin{figure*}[!tpb]
    \centerline{\includegraphics[width=178mm]{spatialattention.eps}}
    \caption{Preference for each position-dependent nucleotide on the sgRNA sequence. 
(a)-(c) Bars show the score of preference after standardization, and the higher the number, the more positive it is for activity of sgRNA. 
The numbers below indicated the position of the nucleotides on-target DNA. 
(d)-(f) Preference surfaces for each position-dependent nucleotide fitted with Bézier surfaces. 
Each position-dependent nucleotide is a control point. 
The coordinates of the control points on the vertical axis represent the degree of preference, 
and the higher the position-dependent nucleotide corresponding to the control point is, the more positive it is for the activity of sgRNA.
The contour plots at the bottom show the area of position-dependent nucleotides which have different contribution to the activity of sgRNA. 
}\label{fig:spatialattention}
\end{figure*}

\subsubsection{Global interpretability}

@@

\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{secondorder.eps}}
    \caption{The visualization of second-order preference matrix $B$, 
the elements in the $i$-th row and the $j$-th column represent the influence of nucleotide at position $j$ when generate the first-order preference $\tilde{A}$ at position $i$. 
The warmer the color, the more important it is. 
In the red box, a few unusual bright spots appear. 
To be more specific, the nucleotide at position 15 has a great effect on the first-order preference at positon 14 and 16.
}\label{fig:heatmap}
\end{figure}

@@

In this paper, we develop AttCRISPR, a ensemble method of both spatial method and temporal method follow the stacking strategy with strong interpretability. 
AttCRISPR prove that the ensemble method have a better performance in the datasets of DeepHF
and can compete with the current state-of-the-art method. 
In addition, AttCRISPR apply attention mechanisms in both the temporal and spatial part, 
and we explain the decisions made by AttCRISPR through the attention module which is consistent with earlier reports. 
Further, we also discovered that the output of the attention module can be used to optimize the low-activity sgRNA without exhaustively search, 
and our optimization results are verified with available experimental data. 
To our knowledge, this is the first time that the attentional mechanim is used to reveal the reasons influencing on-target activity of sgRNA, and assist sgRNA design.

>>>>

\abstract{\textbf{Motivation:} More and more Cas9 variants with higher specificities are developed to avoid the off-target effect, which bring a significant volume of experimental data. 
Conventional machine learning performs poorly on these datasets, while the models based on deep learning are often lack of interpretability, 
which makes researchers have to trade-off accuracy and interpretability. 
We believe it is necessary to develope a model which can not only match deep learning based model in performance, 
but also with strong interpretability that can be comparable to conventional machine learning method.\\
\textbf{Results:} To overcome these problems, we design and implement AttCRISPR, an intrinsic interpretable method based on deep learning to predict the on-target activity. 
In this paper, we find that based on the preprocessing of the input base sequence, the deep learning models can be divided into two categories, 
encoded-based methods and embedded-based methods. 
Based on the categorization above, we assume encoded-based methods and embedded-based methods are heterogeneous, and use a full connection layer to stack them, which achieve a better performance. 
Our model is trained and tested on the DeepHF dataset for performance evaluation. 
AttCRISPR can compete with state-of-the-art method in these datasets, and achieves an average spearman value of 0.872, 0.867, 0.867 (corresponding to WT-SpCas9, eSpCas9(1.1), SpCas9-HF1) with the same evaluation strategy. 
In addition, benefits from two attention modules, spatial and temporal attention modules, AttCRISPR has strong interpretability. 
Through them, we can understand the decisions made by AttCRISPR at both global and local levels without other post hoc explanations techniques. 
With the trained model, we reveal the preference for each position-dependent nucleotide on the sgRNA sequence in each dataset at global level. 
And at local level, we prove that interpretability of AttCRISPR can be used to guide researcher to design sgRNA with higher activity, in order to reduce the experimental cost. 
\\
\textbf{Availability:} The example code are available at \href{https://github.com/South-Walker/AttCRISPR}{https://github.com/South-Walker/AttCRISPR}\\
\textbf{Contact:} \href{zrjiang@cs.ecnu.edu.cn}{zrjiang@cs.ecnu.edu.cn}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
online.}

@@

Clustered regularly interspaced short palindromic repeats (CRISPR) / CRIPSR associated protein 9 (Cas9) systems are preferred over other biological research and human medicine technologies now, 
beacuse of it's efficiency, robustness and programmability. 
Cas9 nucleases can be directed by short guide RNAs (sgRNAs) to introduce site-specific DNA double-stranded breaks in target, 
so to enable editing site-specific regions within genome \citep{jinek2012a,cong2013multiplex,mali2013rnaguided}. 
CRISPR/Cas9, to a large extent, has developed genetic therapies at the cellular level, 
while there are still severe medical disadvantages even now which has greatly hindered the further clinical application of the CRISPR/Cas9 systems. 
One of these disadvantages is due to unexpected insertion and deletion caused by off-target effect \citep{rubeis2018risks,kang2016introducing,ishii2017reproductive,liang2015crispr/cas9-mediated}. 
To overcome this disadvantage, a solution is to engineer CRISPR / Cas9 with higher specificities. 
That's why more and more higher specificities Cas9 variants, such as enhanced SpCas9 (eSpCas9(1.1)), 
Cas9-High Fidelity (SpCas9-HF1) \citep{slaymaker2016rationally,ishii2017reproductive}, hyper-accurate Cas9 (HypaCas9) \citep{kleinstiver2016high-fidelity}, 
been developed and bring a significant volume of experimental data, which means that researchers have to face the difficulty of analyzing such huge and heterogeneous data.

@@

 In addition, the interpretability of existing models are all at global level, and result is a general pattern in the datasets. 
For example, which position has a great impact on the final decision of the model and which position-dependent nucleotide has a positive impact on the activity. 
Lack of analysis at local level. 
For example, which structure causes the low activity of a certain nucleotide sequence and how to improve its activity with a few modification. 
In light of the above, we believe it is necessary to develope a model which can not only match deep learning based model in performance, 
but also with strong interpretability that can be comparable to conventional machine learning method.

@@

Attention mechanism has demonstrated its power in NLP, Statistical Learning, Speech and Computer Vision. 
 It makes model tends to focus selectively on parts of the input, which is help in performing the task effectively. 
 Previous observation have shown that Cas9 preferentially binds sgRNAs containing purines but not pyrimidines \citep{wang2014genetic} and multiple thymine in the spacer impairing sgRNA activity \citep{wu2014genome-wide}, 
 that is to say some specific nucleotides and base position need more attention compared to others. 
 The above is the premise of introducing attention mechanism. Strictly speaking, we are not the first to bring attention mechanisms into this field. 
 The most similar approach to ours is the work based on transformer by \citeauthor{Liu2019}. 
 They use transformer, a components based on attention mechanism, instead of RNN to improve the ability of temporal feature extraction, 
 hence, enhance the performance of their model \citep{vaswani2017attention,Liu2019}. 
 In our work,the interpretability benefit from attention mechanism is more focused. 
 Our main contributions are as follows:\vspace*{1pt}
\begin{itemize}
    \item Present a novel deep-learning model, which can extract potential feature representation of sgRNA sequence in both spatial and temporal domain parallelly. 
    Finally, the ensemble learning method is used to combine the two to achieve better performance than current state-of-the-art models. 
    \item Introduce attention mechanism into our model. 
    As a result, it does not need post hoc explanations techniques based on input perturbation to explain itself. 
    It is intrinsic interpretable in both temporal and spatial domains.
    In the spatial domain it's at global level, while at local level in the temporal domain. 
    Thus, it is transformed from a black box to an intrinsically interpretable model with the performance of deep learning based model. 
    \item Throught ablation analysis and testing a series of possible network structures, 
    we find there are multiple components and strategy can improve the performance of AttCRISPR, 
    which could outperform current state-of-the-art tools on DeepHF dataset.\vspace*{1pt}
\end{itemize}

\section{Materials and methods}
\subsection{Datasets}

The dataset we used for training, validation and testing is built by \citeauthor{wang2019optimized}. 
We extracted 55604, 58617, 56888 sgRNAs with activity (represented by insertion/deletion (indel)) for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively, from its source data.

@@

For encoding process, we use the complementary base to represent the original base in sgRNA. 
Further, we use one-hot encode strategy, that is to say, we encoding each base in sgRNA into a four-dimensional vector 
(encode A,T,G,C into [1,0,0,0], [0,1,0,0], [0,0,1,0], [0,0,0,1], respectively), called one-hot vector. 
Then a sgRNA can be considered as a matrix $X_{oh}\in\mathbb{R}^{l\times4} $, named one-hot matrix (a little sparse, since an one-hot vector is zero in all but one dimension). 
We believe it is meaningful to regard $X_{oh}$ as a binary image, 
therefore, it is used as an input of convolutional neural network, which performs well in the image field.
Meanwhile, as mentioned above, one-hot matrix is a little sparse.

To facilitate the training process, we can mapping each one-hot vector into a dense real-valued high-dimensional space, which is called embedding. 
In summarized, at the matrix level, the formula is as follows
\begin{equation}
X_e=X_{oh}E_m\label{eq:08}
\end{equation}
where $X_e$ named embedding matrix, $E_m\in\mathbb{R}^{4\times m}$ is a trainable transformational matrix, $m$ refers to the dimension of embedding space. 
We believe it is also meaningful to regard nucleotides in the sgRNA sequence as word, and the sgRNA sequence itself as a sentence, 
Guided by this belief $E_m$ is the word embedding matrix and $X_e$ is the sentence embedding in NLP.
Therefore, $X_e$ is used as an input of RNN (or its variant), which performs well in the NLP field.

Due to each element of $X_{oh}$ is interpretable (representing whether there is a corresponding nucleotide type at the corresponding location), 
we call $X_{oh}$ the spatial input, and the CNN work on $X_{oh}$ is method in spatial domain. 
In other hand, different from $X_{oh}$, $X_e$ can only be explained in the first dimension (representing the embedding vector of corresponding nucleotide type), 
and embedding vector is difficult for humans to understand. 
That's why we call $X_e$ the temporal input, and the RNN (or its variant) work on $X_e$ belongs to the method in temporal domain.

@@

In order to introduce the neural network architecture of AttCRISPR, Let's define first-order preference and second-order preference for convenience. 
Taking a simple linear regression model as an example, for input $X$ in Equation~(\ref{eq:01}), $y$ in Equation~(\ref{eq:01}) is as following
 \begin{equation}
y = AX\label{eq:linear}
\end{equation}
where $A \in\mathbb{R}^d$, the total differential of $y$ in Equation~(\ref{eq:linear}) is as following
 \begin{equation}
dy = \sum^{d}_iA_idX_i\label{eq:02}
\end{equation}
where $A_i$ and $X_i$ denotes the $i$-th dimension of the vector $X$ and $A$, 
$A_i$ indicates how dramatically the function changes as $X_i$ changes in a neighborhood of $X$, 
in other word, the importance of $X_i$. 
That's why we'll call $A$ first-order preference in our paper. 
Specifically, we use a vector $A_i$ to build the first-order original preference at position $i$ within sgRNA sequence, 
and $X_i$ is an embeddedness of the $i$-th feature, then $A$ and $X$ are two matrix. 
Further, the final result can be weighted by a trainable non-negative weight vector $W\in\mathbb{R}^l$, as follow
\begin{equation}
y=W \cdot AX^{T}\label{eq:superlinear}
\end{equation}
then we define $\tilde{A}$ as the first-order combine preference matrix (or just first-order preference), which means $\tilde{A}$ can be expressed linearly by $A$ as follow
\begin{equation}
\tilde{A}=BA\label{eq:04}
\end{equation}
where the weight matrix $B\in\mathbb{R}^{l\times l}$ is learned through attention mechanism, 
which we'll call the second-order preference matrix in our paper due to its calculation is based on first-order preference, 
it can explain how a particular pattern containing two nucleotides affects the base sequence. 
Then the predicted value can be expressed as
\begin{equation}
y=W \cdot \tilde{A} {X_e}^{T}\label{eq:05}
\end{equation}

@@

As shown in Figure~\ref{fig:spatialmodule}, given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{array}{l}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{array}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other words, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, we train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{jinek2012a,cong2013multiplex,mali2013rnaguided}.

@@

@@

Then we use a simple full connection network to extract the indirect features, and call the output of fully connection network $y_{bio}$. 
As mentioned above, we assume that the method in spatial domain and in temporal domain can satisfy the diversity premise of ensemble learning.
That's why we follow the stacking strategy, to integrate the methods in time domain and space domain. 
Specifically, the $y_{bio}$, the spatial output $y_{s}$ and the temporal output $y_{t}$ we got earlier are concatenated and then weighted averaging is performed through a full connection layer as follow
\begin{equation}
y=W[y_{bio};y_s;y_t]\label{eq:20}
\end{equation}
where, $y$ is the final prediction value of AttCRISPR, $W$ is the weight learned by the full connection network. 
In the actual implementation, we freeze the network in the spatial domain and temporal domain firstly, in order to make our network focused on learning the weight $W$.
Then the parameters of the entire network are adjusted in the fine tuning of AttCRISPR.

\subsection{Current prediction method}

As far as we know, DeepHF, the state-of-the-art method, 
achieved Spearman correlation coefficients of 0.867, 0.862 and 0.860 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively \citep{wang2019optimized}. 
Another tool worth noting is CRISPRpred(SEQ), which is based on conventional machine learning and need complicated human feature engineering exercise \citep{MuhammadRafid2020}. 
They also test their tool with this dataset and achieved Spearman correlation coefficients of 0.838, 0.830 and 0.821 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively (without any hyperparameter tuning).

In order to make the comparison apples to apples, we follow the same strategy as \citeauthor{wang2019optimized} to design our experiments.
To be more specific, each set is shuffled and divided into three parts, 76.5\%, 15\% and 8.5\% of the relevant data was used as the training, test and validation set respectively in a single experiment. 
The experiment is repeated ten times with the results recorded and averaged finally.

@@

\section{Results}
\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{baselinewithoutbiofeat.eps}}
    \caption{In the absence of hand-crafted biological features, performance of different algorithms for sgRNA activity prediction. 
    (a)-(c) The performance of Temporal AttCRISPR, Spatial AttCRISPR and Ensemble AttCRISPR. 
    The half-violin plots show the mean and distribution of the Spearman correlation coefficient between predicted and measured sgRNA activity scores over all tests. 
    (d)-(f) In the absence of hand-crafted biological features, the performance of all prediction methods in these datasets as far as we know. 
    The $mean \pm s.d.$ of the Spearman correlation coefficient between predicted and measured sgRNA activity scores are shown in the bar plots.}\label{fig:06}
\end{figure}

We designed experiments to address the following questions:
\begin{itemize}
    \item 
    In the absence of hand-crafted biological features, whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions made by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

\begin{figure*}[!tpb]
    \centerline{\includegraphics[width=178mm]{spatialattention.eps}}
    \caption{Preference for each position-dependent nucleotide on the sgRNA sequence. 
(a)-(c) Bars show the score of preference after standardization, and the higher the number, the more positive it is for activity of sgRNA. 
The numbers below indicated the position of the nucleotides on-target DNA. 
(d)-(f) Preference surfaces for each position-dependent nucleotide fitted with Bézier surfaces. 
Each position-dependent nucleotide is a control point. 
The coordinates of the control points on the vertical axis represent the degree of preference, 
and the higher the position-dependent nucleotide corresponding to the control point is, the more positive it is for the activity of sgRNA.
The contour plots at the bottom show the area of position-dependent nucleotides which have different contribution to the activity of sgRNA. 
}\label{fig:spatialattention}
\end{figure*}

@@

In the following sections, we will analyze the insight into activity of sgRNA bring through the attention mechanism at both global and local levels 
to validate the attention module in the AttCRISPR can help us to understand the decisions. 
\subsubsection{Global interpretability}

@@

@@

\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{secondorder.eps}}
    \caption{The visualization of second-order preference matrix $B$, 
the elements in the $i$-th row and the $j$-th column represent the influence of nucleotide at position $j$ when generate the first-order preference $\tilde{A}$ at position $i$. 
The warmer the color, the more important it is. 
In the red box, a few unusual bright spots appear. 
To be more specific, the nucleotide at position 15 has a great effect on the first-order preference at positon 14 and 16.
}\label{fig:heatmap}
\end{figure}
In this paper, we develop AttCRISPR, a ensemble method of both spatial method and temporal method follow the stacking strategy with strong interpretability. 
AttCRISPR prove that the ensemble method have a better performance in the datasets of DeepHF
and can compete with the current state-of-the-art method. 
In addition, AttCRISPR apply attention mechanisms in both the temporal and spatial part, 
and we explain the decisions made by AttCRISPR through the attention module which is consistent with earlier reports. 
Further, we also discovered that the output of the attention module can be used to optimize the low-activity sgRNA without exhaustively search, 
and our optimization results are verified with available experimental data. 
To our knowledge, this is the first time that the attentional mechanim is used to reveal the reasons influencing on-target activity of sgRNA, and assist sgRNA design.
